---
title: "Visualising spatial data tutorial"
author: "Eric Roseren"
date: "2/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages for spatial data

A range of R packages for spatial data can be found at <http://cran.r-project.org/web/views/Spatial.html>. 

In this tutorial we will use the packages from the ‘spverse’, that use the sp package:
*ggmap
*rgdal
*rgeos
*maptools
*tidyverse
*tmap

```{r echo=FALSE Installing-pakages}
packages <- c("ggmap", "rgdal", "rgeos", "maptools", "dplyr", "tidyr", "tmap")
# install.packages(x) # warning: uncommenting this may take a number of minutes)
lapply(packages, library, character.only = TRUE) # load the required packages
```

The data used for this tutorial can be downloaded from <https://github.com/Robinlovelace/Creating-maps-in-R>. 
```{r echo=FALSE Loading-data}
lnd <- readOGR(dsn = "data/london_sport.shp")
```

## Exploratory Data Analysis

```{r}
head(lnd@data, n = 2)
mean(lnd$Partic_Per)

lnd$Pop_2001 <- as.numeric(as.character(lnd$Pop_2001))
nrow(lnd);ncol(lnd)
```
## Plots

```{r}
plot(lnd)
plot(lnd@data)
```
select rows of lnd@data where sports participation is less than 13:
```{r}
lnd@data[lnd$Partic_Per<13,]
# or 
lnd@data %>% filter(Partic_Per<13)
```

Select zones where sports participation is between 20 and 25%
```{r}
sel <- which(lnd$Partic_Per > 20 & lnd$Partic_Per < 25)
plot(lnd[sel, ]) # output not shown here
head(sel)

plot(lnd, col = "lightgrey") # plot the london_sport object
sel <- lnd$Partic_Per > 25
plot(lnd[ sel, ], col = "turquoise", add = TRUE) # add selected zones to map
```

## Challenge
Select all zones whose geographic centroid lies within 10 km of the geographic
centroid of inner London.
```{r}
centroids <- gCentroid(lnd,byid = TRUE)
dist.matrix <- as.data.frame(gDistance(centroids,byid = T))
center.lnd <- gCentroid(lnd[lnd$name == "City of London",]) 

```
Get centroid from polygon using the gCentroid function from rgeos`

```{r}
crnames<-as.character(lnd@data$name)
colnames(dist.matrix)<- crnames
rownames(dist.matrix)<- crnames

dist.to.london.ind <- which(dist.matrix$`City of London`<=10000)  
```
## Plot of London geographic centroid area within 10km of city centre
```{r Circle function}
x <- centroids@coords[33,][1]
y <- centroids@coords[33,][2]
t=seq(0,2*pi,length=200)
r <- 10000
```

```{r}
t=seq(0,2*pi,length=200)
plot(lnd, col = "lightgrey") # plot the london_sport object
plot(lnd[ dist.to.london.ind, ], col = "turquoise",border='grey', add = TRUE)
lines(x+r*cos(t),y+r*sin(t),col='red')# adding circle of 10km
text(coordinates(center.lnd), "Central\nLondon")
```

## Solutions for challenge 1:
```{r}
plot(lnd, col = "grey")
# find London's geographic centroid (add ", byid = T" for all)
cent_lnd <- gCentroid(lnd[lnd$name == "City of London",]) 
points(cent_lnd, cex = 3)
# set 10 km buffer
lnd_buffer <- gBuffer(spgeom = cent_lnd, width = 10000,capStyle = 'ROUND') 


# method 1 of subsetting selects any intersecting zones
lnd_central <- lnd[lnd_buffer,] # the selection is too big!
# test the selection for the previous method - uncomment below
plot(lnd_central, col = "lightblue", add = T)
plot(lnd_buffer, add = T) # some areas just touch the buffer

# method2 of subsetting selects only points within the buffer
lnd_cents <- SpatialPoints(coordinates(lnd),
  proj4string = CRS(proj4string(lnd))) # create spatialpoints
sel <- lnd_cents[lnd_buffer,] # select points inside buffer
points(sel) # show where the points are located
lnd_central <- lnd[sel,] # select zones intersecting w. sel
plot(lnd_central, add = T, col = "lightslateblue", 
  border = "grey")
plot(lnd_buffer, add = T, border = "red", lwd = 2)
# Add text to the plot!
text(coordinates(cent_lnd), "Central\nLondon")
```


## Selecting quadrant

The 4 quadrants of London and dissolved borders. Challenge 2: Recreate the plot displayed in the tutorial:
```{r}
# Find the centre of the london area
easting.lnd <- coordinates(gCentroid(lnd))[[1]]
northing.lnd <- coordinates(gCentroid(lnd))[[2]]
```
```{r}
# arguments to test whether or not a coordinate is east or north of the centre
east <- sapply(coordinates(lnd)[,1], function(x) x > easting.lnd)
north <- sapply(coordinates(lnd)[,2], function(x) x > northing.lnd)
west <- sapply(coordinates(lnd)[,1], function(x) x < easting.lnd)
south <- sapply(coordinates(lnd)[,2], function(x) x < northing.lnd)
```
```{r}
# test if the coordinate is east and north of the centre
lnd$quadrant <- "unknown" # prevent NAs in result
lnd$quadrant[east & north] <- "northeast"
lnd$quadrant[east & south] <- "southeast"
lnd$quadrant[west & north] <- "northwest"
lnd$quadrant[west & south] <- "southwest"
```
Plot of the four quadrant:
```{r}
sel.NE <- which(lnd$quadrant=="northeast")
plot(lnd[sel.NE, ]) # output not shown here
head(sel.NE)

sel.NW <- which(lnd$quadrant=="northwest")
plot(lnd[sel.NW, ]) # output not shown here
head(sel.NW)

sel.SE <- which(lnd$quadrant=="southeast")
plot(lnd[sel.SE, ]) # output not shown here
head(sel.SE)

sel.SW <- which(lnd$quadrant=="southwest")
plot(lnd[sel.SW, ]) # output not shown here
head(sel.SW)

plot(lnd, col = "lightgrey") # plot the london_sport object
plot(lnd[ sel.NE, ], col = "turquoise", add = TRUE) # add selected zones to map
plot(lnd[ sel.NW, ], col = "yellow", add = TRUE) # add selected zones to map
plot(lnd[ sel.SE, ], col = "purple", add = TRUE) 
plot(lnd[ sel.SW, ], col = "red", add = TRUE) 
llgridlines(lnd)
```
## Solution of the second challenge:

```{r}
plot(lnd)
plot(lnd[east & north,], add = TRUE, col = "red" )
llgridlines(lnd, lty= 3, side ="EN", offset = -0.5)

lnd_disolved = rgeos::gUnaryUnion(spgeom = lnd, id = lnd$quadrant)

qtm(lnd, fill = "quadrant") +
  tm_shape(lnd_disolved) +
  tm_borders(lwd = 9)
```

# Part III: Creating and manipulating spatial data

```{r}
vec <- vector(mode = "numeric", length = 3)
df <- data.frame(x = 1:3, y = c(1/2, 2/3, 3/4))

sp1 <- SpatialPoints(coords = df)

EPSG <- make_EPSG() # create data frame of available EPSG codes
EPSG[grepl("WGS 84$", EPSG$note), ] # search for WGS 84 code
lnd84 <- spTransform(lnd, CRS("+init=epsg:4326")) # reproject
# Save lnd84 object (we will use it in Part IV)
saveRDS(object = lnd84, file = "data/lnd84.Rds")

```
##  Linking additional pieces of information to our polygons.

```{r}
# Create new object called "lnd" from "london_sport" shapefile
lnd <- readOGR("data/london_sport.shp")
plot(lnd) # plot the lnd object (not shown)
nrow(lnd) 
```

```{r}
# Create and look at new crime_data object
crime_data <- read.csv("data/mps-recordedcrime-borough.csv",
stringsAsFactors = FALSE)
head(crime_data$CrimeType) # information about crime type
```

```{r}
crime_data %>% filter(CrimeType=="Theft & Handling") -> crime_theft
```
Calculate the sum of the crime count for each district, save result

```{r}
crime_ag <- crime_theft %>% group_by(Borough) %>% summarise(crime.number = sum(CrimeCount)) 
crime.ag <- aggregate(CrimeCount ~ Borough, FUN = sum, data = crime_theft)
```
```{r}
# Compare the name column in lnd to Borough column in crime_ag to see which rows match.
lnd$name %in% crime.ag$Borough
# Return rows which do not match
lnd$name[!lnd$name %in% crime_ag$Borough]
crime_ag$Borough[!crime_ag$Borough %in% lnd$name]
```
Join the datasets
```{r}
# head(left_join(lnd@data, crime_ag)) # test it works
lnd@data <- left_join(lnd@data, crime.ag, by = c('name' = 'Borough'))
```

Plot the rate of theft crimes in London by borough.

```{r}
qtm(lnd, "CrimeCount")
```

Optional challenges:

```{r}
library(readxl)
rent.price <- read_xls("data/average-rent-borough.xls", sheet = 2)
head(rent.price) # information about crime type
```











